---
title: Loanword f adopted as labialized x
author: Nikita Beklemishev
date: 'Last update: `r lubridate::make_datetime(year = 2023, month = 4, day = 7)`'
output:
  html_document:
    number_sections: true
    anchor_sections: true
    pandoc_args: --shift-heading-level-by=-1
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE, fig.width = 9.5)
library(tidyverse)
library(lingtypology)
read_csv('data/database.csv', show_col_types = FALSE) |> 
  filter(feature_id == 16) ->
  db
read_csv('data/villages.csv') |> 
  filter(!(village %in% c('Kazankulak', 'Novyy Borch', 'Vrush', 'Aran', 'Khnyukh'))) ->
  villages
  
villages |> 
  filter(!(village %in% c('Tsudik', 'Borch'))) ->
  villages4map
  
```

Many Nakh-Daghestan languages lack /f/ in their phonemic inventories. [Chirikba 2008: 50] When borrowing a loanword with /f/, like Azeri tüfängʲ 'rifle' or Arabic fikr 'thought', loan /f/ is usually rendered as /p/ or /b/ (see St. Dargwa tupang, pikri). Likewise, for Rutul, Ibragimov (2004: 32) reports that in most dialects /f/ is imported as /xʷ/ (which is acoutically similar and native to Rutul) or /p/: "xʷatima < Ar. faːtˁima, pindiqʼ < Az. fındıq". This paticular feature indicates whether /xʷ/ can appear instead of the loanword /f/. In practice, we diagnose this by asking for inter-dialectal lexemes that supposedly have arabic-turkic-persian origin, containing sound /f/ in source languages. The sounds in check (preliminary values) are the sounds occuring in place of /f/ in words /aːferin/ 'acclamation, respect', /kef/ 'pleasure, relax, keif', /findɨqʼ/ 'hazlenut', /χˁafte/ 'a week', /χilaf/ 'lie, tratory'. In final values, sounds like [xʷ] or [ф] would be labelled as <yes> (e.g. source /f/ underwent adaptation), [f] is labelled as <no> (no adaptation), and when both sounds occured (e.g. maybe they are not contrastive) I used the label <variation>.


## “acclamation, respect” //aːferin// {.tabset .tabset-fade .tabset-pills}

### Map

```{r}
db |> 
  filter(feature_lexeme == '“acclamation, respect” //aːferin//')  |>
  filter(!is.na(value),
         value != 'NO DATA') |>
  mutate(value = str_split(value, ' ; ')) |> 
  unnest_longer(value) |> 
  distinct(settlement, value) |> 
  mutate(n = 1) |> 
  pivot_wider(names_from = value, values_from = n, values_fill = 0) |> 
  left_join(villages[,c('village', 'lat', 'lon')], c('settlement' = 'village')) |> 
  mutate(language = 'Rutul') ->
  for_map
  
if(length(for_map) == 5){
map.feature(languages = 'Rutul',
            latitude = villages4map$lat,
            longitude = villages4map$lon,
            label = villages4map$village,
            label.position = 'top',
            label.hide = FALSE,
            width = 10,
            color = 'gray',
            tile = 'OpenStreetMap.HOT',
            opacity = 0.4) |> 
  map.feature(languages = for_map$language,
              latitude = for_map$lat,
              longitude = for_map$lon,
              label = for_map$settlement,
              label.position = 'top',
              label.hide = FALSE,
              width = 10,
              tile = 'OpenStreetMap.HOT',
              features = colnames(for_map)[2],
              pipe.data = _)  
} else {
map.feature(languages = 'Rutul',
            latitude = villages4map$lat,
            longitude = villages4map$lon,
            label = villages4map$village,
            label.position = 'top',
            label.hide = FALSE,
            width = 10,
            color = 'gray',
            tile = 'OpenStreetMap.HOT',
            opacity = 0.4) |> 
  map.feature(languages = for_map$language,
            latitude = for_map$lat,
            longitude = for_map$lon,
            minichart.data = for_map |> select(-settlement, -lat, -lon, -language),
            minichart = 'pie', 
            width = 3,
            tile = 'OpenStreetMap.HOT',
            pipe.data = _)
}
```

### Data

```{r}
db |> 
  filter(feature_lexeme == '“acclamation, respect” //aːferin//') |> 
  select(settlement, value, stimuli, answer, collected) |> 
  DT::datatable(class = 'cell-border stripe', 
    rownames = FALSE, 
    filter = 'top', 
    extensions = 'Buttons',
    options = list(pageLength = 42, 
                   autoWidth = TRUE, 
                   info = FALSE,
                   dom = 'fBltp',
                   buttons = list(list(extend = 'collection',
                                       buttons = c('csv', 'excel', 'pdf'),
                                       text = '<i class="fas fa-download"></i>')),
                   paginate = TRUE))
```


## “pleasure” //ke(j)F// {.tabset .tabset-fade .tabset-pills}

### Map

```{r}
db |> 
  filter(feature_lexeme == '“pleasure” //ke(j)F//')  |>
  filter(!is.na(value),
         value != 'NO DATA') |>
  mutate(value = str_split(value, ' ; ')) |> 
  unnest_longer(value) |> 
  distinct(settlement, value) |> 
  mutate(n = 1) |> 
  pivot_wider(names_from = value, values_from = n, values_fill = 0) |> 
  left_join(villages[,c('village', 'lat', 'lon')], c('settlement' = 'village')) |> 
  mutate(language = 'Rutul') ->
  for_map
  
if(length(for_map) == 5){
map.feature(languages = 'Rutul',
            latitude = villages4map$lat,
            longitude = villages4map$lon,
            label = villages4map$village,
            label.position = 'top',
            label.hide = FALSE,
            width = 10,
            color = 'gray',
            tile = 'OpenStreetMap.HOT',
            opacity = 0.4) |> 
  map.feature(languages = for_map$language,
              latitude = for_map$lat,
              longitude = for_map$lon,
              label = for_map$settlement,
              label.position = 'top',
              label.hide = FALSE,
              width = 10,
              tile = 'OpenStreetMap.HOT',
              features = colnames(for_map)[2],
              pipe.data = _)  
} else {
map.feature(languages = 'Rutul',
            latitude = villages4map$lat,
            longitude = villages4map$lon,
            label = villages4map$village,
            label.position = 'top',
            label.hide = FALSE,
            width = 10,
            color = 'gray',
            tile = 'OpenStreetMap.HOT',
            opacity = 0.4) |> 
  map.feature(languages = for_map$language,
            latitude = for_map$lat,
            longitude = for_map$lon,
            minichart.data = for_map |> select(-settlement, -lat, -lon, -language),
            minichart = 'pie', 
            width = 3,
            tile = 'OpenStreetMap.HOT',
            pipe.data = _)
}
```

### Data

```{r}
db |> 
  filter(feature_lexeme == '“pleasure” //ke(j)F//') |> 
  select(settlement, value, stimuli, answer, collected) |> 
  DT::datatable(class = 'cell-border stripe', 
    rownames = FALSE, 
    filter = 'top', 
    extensions = 'Buttons',
    options = list(pageLength = 42, 
                   autoWidth = TRUE, 
                   info = FALSE,
                   dom = 'fBltp',
                   buttons = list(list(extend = 'collection',
                                       buttons = c('csv', 'excel', 'pdf'),
                                       text = '<i class="fas fa-download"></i>')),
                   paginate = TRUE))
```


## “hazlenut” //Findɨqʼ// {.tabset .tabset-fade .tabset-pills}

### Map

```{r}
db |> 
  filter(feature_lexeme == '“hazlenut” //Findɨqʼ//')  |>
  filter(!is.na(value),
         value != 'NO DATA') |>
  mutate(value = str_split(value, ' ; ')) |> 
  unnest_longer(value) |> 
  distinct(settlement, value) |> 
  mutate(n = 1) |> 
  pivot_wider(names_from = value, values_from = n, values_fill = 0) |> 
  left_join(villages[,c('village', 'lat', 'lon')], c('settlement' = 'village')) |> 
  mutate(language = 'Rutul') ->
  for_map
  
if(length(for_map) == 5){
map.feature(languages = 'Rutul',
            latitude = villages4map$lat,
            longitude = villages4map$lon,
            label = villages4map$village,
            label.position = 'top',
            label.hide = FALSE,
            width = 10,
            color = 'gray',
            tile = 'OpenStreetMap.HOT',
            opacity = 0.4) |> 
  map.feature(languages = for_map$language,
              latitude = for_map$lat,
              longitude = for_map$lon,
              label = for_map$settlement,
              label.position = 'top',
              label.hide = FALSE,
              width = 10,
              tile = 'OpenStreetMap.HOT',
              features = colnames(for_map)[2],
              pipe.data = _)  
} else {
map.feature(languages = 'Rutul',
            latitude = villages4map$lat,
            longitude = villages4map$lon,
            label = villages4map$village,
            label.position = 'top',
            label.hide = FALSE,
            width = 10,
            color = 'gray',
            tile = 'OpenStreetMap.HOT',
            opacity = 0.4) |> 
  map.feature(languages = for_map$language,
            latitude = for_map$lat,
            longitude = for_map$lon,
            minichart.data = for_map |> select(-settlement, -lat, -lon, -language),
            minichart = 'pie', 
            width = 3,
            tile = 'OpenStreetMap.HOT',
            pipe.data = _)
}
```

### Data

```{r}
db |> 
  filter(feature_lexeme == '“hazlenut” //Findɨqʼ//') |> 
  select(settlement, value, stimuli, answer, collected) |> 
  DT::datatable(class = 'cell-border stripe', 
    rownames = FALSE, 
    filter = 'top', 
    extensions = 'Buttons',
    options = list(pageLength = 42, 
                   autoWidth = TRUE, 
                   info = FALSE,
                   dom = 'fBltp',
                   buttons = list(list(extend = 'collection',
                                       buttons = c('csv', 'excel', 'pdf'),
                                       text = '<i class="fas fa-download"></i>')),
                   paginate = TRUE))
```


## “week” //χˁaFte// {.tabset .tabset-fade .tabset-pills}

### Map

```{r}
db |> 
  filter(feature_lexeme == '“week” //χˁaFte//')  |>
  filter(!is.na(value),
         value != 'NO DATA') |>
  mutate(value = str_split(value, ' ; ')) |> 
  unnest_longer(value) |> 
  distinct(settlement, value) |> 
  mutate(n = 1) |> 
  pivot_wider(names_from = value, values_from = n, values_fill = 0) |> 
  left_join(villages[,c('village', 'lat', 'lon')], c('settlement' = 'village')) |> 
  mutate(language = 'Rutul') ->
  for_map
  
if(length(for_map) == 5){
map.feature(languages = 'Rutul',
            latitude = villages4map$lat,
            longitude = villages4map$lon,
            label = villages4map$village,
            label.position = 'top',
            label.hide = FALSE,
            width = 10,
            color = 'gray',
            tile = 'OpenStreetMap.HOT',
            opacity = 0.4) |> 
  map.feature(languages = for_map$language,
              latitude = for_map$lat,
              longitude = for_map$lon,
              label = for_map$settlement,
              label.position = 'top',
              label.hide = FALSE,
              width = 10,
              tile = 'OpenStreetMap.HOT',
              features = colnames(for_map)[2],
              pipe.data = _)  
} else {
map.feature(languages = 'Rutul',
            latitude = villages4map$lat,
            longitude = villages4map$lon,
            label = villages4map$village,
            label.position = 'top',
            label.hide = FALSE,
            width = 10,
            color = 'gray',
            tile = 'OpenStreetMap.HOT',
            opacity = 0.4) |> 
  map.feature(languages = for_map$language,
            latitude = for_map$lat,
            longitude = for_map$lon,
            minichart.data = for_map |> select(-settlement, -lat, -lon, -language),
            minichart = 'pie', 
            width = 3,
            tile = 'OpenStreetMap.HOT',
            pipe.data = _)
}
```

### Data

```{r}
db |> 
  filter(feature_lexeme == '“week” //χˁaFte//') |> 
  select(settlement, value, stimuli, answer, collected) |> 
  DT::datatable(class = 'cell-border stripe', 
    rownames = FALSE, 
    filter = 'top', 
    extensions = 'Buttons',
    options = list(pageLength = 42, 
                   autoWidth = TRUE, 
                   info = FALSE,
                   dom = 'fBltp',
                   buttons = list(list(extend = 'collection',
                                       buttons = c('csv', 'excel', 'pdf'),
                                       text = '<i class="fas fa-download"></i>')),
                   paginate = TRUE))
```


## “lie, tratory” //χilaf// {.tabset .tabset-fade .tabset-pills}

### Map

```{r}
db |> 
  filter(feature_lexeme == '“lie, tratory” //χilaf//')  |>
  filter(!is.na(value),
         value != 'NO DATA') |>
  mutate(value = str_split(value, ' ; ')) |> 
  unnest_longer(value) |> 
  distinct(settlement, value) |> 
  mutate(n = 1) |> 
  pivot_wider(names_from = value, values_from = n, values_fill = 0) |> 
  left_join(villages[,c('village', 'lat', 'lon')], c('settlement' = 'village')) |> 
  mutate(language = 'Rutul') ->
  for_map
  
if(length(for_map) == 5){
map.feature(languages = 'Rutul',
            latitude = villages4map$lat,
            longitude = villages4map$lon,
            label = villages4map$village,
            label.position = 'top',
            label.hide = FALSE,
            width = 10,
            color = 'gray',
            tile = 'OpenStreetMap.HOT',
            opacity = 0.4) |> 
  map.feature(languages = for_map$language,
              latitude = for_map$lat,
              longitude = for_map$lon,
              label = for_map$settlement,
              label.position = 'top',
              label.hide = FALSE,
              width = 10,
              tile = 'OpenStreetMap.HOT',
              features = colnames(for_map)[2],
              pipe.data = _)  
} else {
map.feature(languages = 'Rutul',
            latitude = villages4map$lat,
            longitude = villages4map$lon,
            label = villages4map$village,
            label.position = 'top',
            label.hide = FALSE,
            width = 10,
            color = 'gray',
            tile = 'OpenStreetMap.HOT',
            opacity = 0.4) |> 
  map.feature(languages = for_map$language,
            latitude = for_map$lat,
            longitude = for_map$lon,
            minichart.data = for_map |> select(-settlement, -lat, -lon, -language),
            minichart = 'pie', 
            width = 3,
            tile = 'OpenStreetMap.HOT',
            pipe.data = _)
}
```

### Data

```{r}
db |> 
  filter(feature_lexeme == '“lie, tratory” //χilaf//') |> 
  select(settlement, value, stimuli, answer, collected) |> 
  DT::datatable(class = 'cell-border stripe', 
    rownames = FALSE, 
    filter = 'top', 
    extensions = 'Buttons',
    options = list(pageLength = 42, 
                   autoWidth = TRUE, 
                   info = FALSE,
                   dom = 'fBltp',
                   buttons = list(list(extend = 'collection',
                                       buttons = c('csv', 'excel', 'pdf'),
                                       text = '<i class="fas fa-download"></i>')),
                   paginate = TRUE))
```


