---
title: Paradigm levelling of šu “brother”
author: Nikita Beklemishev
date: 'Last update: `r lubridate::make_datetime(year = 2023, month = 6, day = 24)`'
output:
  html_document:
    number_sections: false
    anchor_sections: true
    pandoc_args: --shift-heading-level-by=-1
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE, fig.width = 9.5)
library(tidyverse)
library(lingtypology)
read_csv('data/database.csv', show_col_types = FALSE) |> 
  filter(feature_id == 43) ->
  db
read_csv('data/villages.csv') |> 
  filter(!(village %in% c('Kazankulak', 'Novyy Borch', 'Vrush', 'Aran', 'Khnyukh'))) ->
  villages
  
villages |> 
  filter(!(village %in% c('Tsudik', 'Borch'))) ->
  villages4map
  
```

Umlautization of the stem, e.g. applying the [+front, −back] features to the stem vowels, as a means of marking the PL forms appears not only in verbal stems (see phonology feature [@Vowel fronting in APL]). Something of similar nature is found in paradigm of the lexeme /šu/ ‘brother’ and a few other kin terms in some dialects. While the regular formation would be adding a -bɨr/-mar suffix, actually the suffix -bä (-be, bäˁ, bâ) is added. Moreover, the noun takes a special plural stem /šü/. [@filatovms §8.1 @danielfilatovms §1.3] This feature indicates whether the stem paradigm levells (and by plural or by singular form) or the alternation remains. Preliminary values are vowels in the first syllable of ‘brother.SG’ and the vowels of the first syllabe of ‘brother.PL’. The values u / uˁ in singular with ü / üˁ in plural are grouped together as **alternating** stem, values u / uˁ in both singular and plural are grouped as levelled by **SG stem**. Values where ü / üˁ occurs in both singular and plural are grouped as levelled by **PL stem**. Cases where ‘brother’ forms plural with the reglar affix are marked as **no irregularity**.


## ‘brother’ /šu/, ‘brothers’ /šübe/

::: {.panel-tabset}

### Map

```{r}
db |> 
  filter(feature_lexeme == '‘brother’ /šu/, ‘brothers’ /šübe/')  |>
  filter(!is.na(value),
         value != 'NO DATA') |>
  mutate(value = str_split(value, ' ; ')) |> 
  unnest_longer(value) |> 
  distinct(settlement, value) |> 
  mutate(n = 1) |> 
  pivot_wider(names_from = value, values_from = n, values_fill = 0) |> 
  left_join(villages[,c('village', 'lat', 'lon')], c('settlement' = 'village')) |> 
  mutate(language = 'Rutul') ->
  for_map
  
if(length(for_map) == 5){
map.feature(languages = 'Rutul',
            latitude = villages4map$lat,
            longitude = villages4map$lon,
            label = villages4map$village,
            label.position = 'top',
            label.hide = FALSE,
            width = 10,
            color = 'gray',
            tile = 'OpenStreetMap.HOT',
            opacity = 0.4) |> 
  map.feature(languages = for_map$language,
              latitude = for_map$lat,
              longitude = for_map$lon,
              label = for_map$settlement,
              label.position = 'top',
              label.hide = FALSE,
              width = 10,
              tile = 'OpenStreetMap.HOT',
              features = colnames(for_map)[2],
              pipe.data = _)  
} else {
map.feature(languages = 'Rutul',
            latitude = villages4map$lat,
            longitude = villages4map$lon,
            label = villages4map$village,
            label.position = 'top',
            label.hide = FALSE,
            width = 10,
            color = 'gray',
            tile = 'OpenStreetMap.HOT',
            opacity = 0.4) |> 
  map.feature(languages = for_map$language,
            latitude = for_map$lat,
            longitude = for_map$lon,
            minichart.data = for_map |> select(-settlement, -lat, -lon, -language),
            minichart = 'pie', 
            width = 3,
            tile = 'OpenStreetMap.HOT',
            pipe.data = _)
}
```

### Data

```{r}
db |> 
  filter(feature_lexeme == '‘brother’ /šu/, ‘brothers’ /šübe/') |> 
  select(settlement, value, stimuli, answer, collected) |> 
  DT::datatable(class = 'cell-border stripe', 
    rownames = FALSE, 
    filter = 'top', 
    extensions = 'Buttons',
    options = list(pageLength = 42, 
                   autoWidth = TRUE, 
                   info = FALSE,
                   dom = 'fBltp',
                   buttons = list(list(extend = 'collection',
                                       buttons = c('csv', 'excel', 'pdf'),
                                       text = '<i class="fas fa-download"></i>')),
                   paginate = TRUE))
```
:::

