---
title: The verb ‘say’, perfective and imperfective stems
author: Timur Maisak
date: 'Last update: `r lubridate::make_datetime(year = 2023, month = 12, day = 5)`'
output:
  html_document:
    number_sections: true
    anchor_sections: true
    pandoc_args: --shift-heading-level-by=-1
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE, fig.width = 9.5)
library(tidyverse)
library(lingtypology)
read_csv('data/database.csv', show_col_types = FALSE) |> 
  filter(feature_id == 93) ->
  db
read_csv('data/villages.csv') |> 
  filter(!(village %in% c('Kazankulak', 'Novyy Borch', 'Vrush', 'Aran', 'Khnyukh'))) ->
  villages
  
villages |> 
  filter(!(village %in% c('Tsudik', 'Borch'))) ->
  villages4map
  
```

The main speech verb 'say' is a high-frequency verb in Rutul, with cross-dialectal variation in the forms of aspectual stems. For the perfective stem, the stimulus sentence was 'He told the truth'. In the Rutul translation, one of the forms based on the perfective stem (usually, the aorist) was used. For the imperfective stem, the stimulus sentence was 'He always says "yes"'. In the Rutul translation, one of the periphrastic present tense forms based on the imperfective converb or participle was used.
As the map value, only the stems are given, without converb or participle suffixes. The gender is unmarked in this stem (in the "weak" gender agreement series, Gender 4 has a zero affix).


## Imperfective stem {.tabset .tabset-fade .tabset-pills}

### Map

```{r}
db |> 
  filter(feature_lexeme == 'Imperfective stem')  |>
  filter(!is.na(value),
         value != 'NO DATA') |>
  mutate(value = str_split(value, ' ; ')) |> 
  unnest_longer(value) |> 
  distinct(settlement, value) |> 
  mutate(n = 1) |> 
  pivot_wider(names_from = value, values_from = n, values_fill = 0) |> 
  left_join(villages[,c('village', 'lat', 'lon')], c('settlement' = 'village')) |> 
  mutate(language = 'Rutul') ->
  for_map
  
if(length(for_map) == 5){
map.feature(languages = 'Rutul',
            latitude = villages4map$lat,
            longitude = villages4map$lon,
            label = villages4map$village,
            label.position = 'top',
            label.hide = FALSE,
            width = 10,
            color = 'gray',
            tile = 'OpenStreetMap.HOT',
            opacity = 0.4) |> 
  map.feature(languages = for_map$language,
              latitude = for_map$lat,
              longitude = for_map$lon,
              label = for_map$settlement,
              label.position = 'top',
              label.hide = FALSE,
              width = 10,
              tile = 'OpenStreetMap.HOT',
              features = colnames(for_map)[2],
              pipe.data = _)  
} else {
map.feature(languages = 'Rutul',
            latitude = villages4map$lat,
            longitude = villages4map$lon,
            label = villages4map$village,
            label.position = 'top',
            label.hide = FALSE,
            width = 10,
            color = 'gray',
            tile = 'OpenStreetMap.HOT',
            opacity = 0.4) |> 
  map.feature(languages = for_map$language,
            latitude = for_map$lat,
            longitude = for_map$lon,
            minichart.data = for_map |> select(-settlement, -lat, -lon, -language),
            minichart = 'pie', 
            width = 3,
            tile = 'OpenStreetMap.HOT',
            pipe.data = _)
}
```

### Data

```{r}
db |> 
  filter(feature_lexeme == 'Imperfective stem') |> 
  select(settlement, value, stimuli, answer, collected) |> 
  DT::datatable(class = 'cell-border stripe', 
    rownames = FALSE, 
    filter = 'top', 
    extensions = 'Buttons',
    options = list(pageLength = 42, 
                   autoWidth = TRUE, 
                   info = FALSE,
                   dom = 'fBltp',
                   buttons = list(list(extend = 'collection',
                                       buttons = c('csv', 'excel', 'pdf'),
                                       text = '<i class="fas fa-download"></i>')),
                   paginate = TRUE))
```


## Perfective stem {.tabset .tabset-fade .tabset-pills}

### Map

```{r}
db |> 
  filter(feature_lexeme == 'Perfective stem')  |>
  filter(!is.na(value),
         value != 'NO DATA') |>
  mutate(value = str_split(value, ' ; ')) |> 
  unnest_longer(value) |> 
  distinct(settlement, value) |> 
  mutate(n = 1) |> 
  pivot_wider(names_from = value, values_from = n, values_fill = 0) |> 
  left_join(villages[,c('village', 'lat', 'lon')], c('settlement' = 'village')) |> 
  mutate(language = 'Rutul') ->
  for_map
  
if(length(for_map) == 5){
map.feature(languages = 'Rutul',
            latitude = villages4map$lat,
            longitude = villages4map$lon,
            label = villages4map$village,
            label.position = 'top',
            label.hide = FALSE,
            width = 10,
            color = 'gray',
            tile = 'OpenStreetMap.HOT',
            opacity = 0.4) |> 
  map.feature(languages = for_map$language,
              latitude = for_map$lat,
              longitude = for_map$lon,
              label = for_map$settlement,
              label.position = 'top',
              label.hide = FALSE,
              width = 10,
              tile = 'OpenStreetMap.HOT',
              features = colnames(for_map)[2],
              pipe.data = _)  
} else {
map.feature(languages = 'Rutul',
            latitude = villages4map$lat,
            longitude = villages4map$lon,
            label = villages4map$village,
            label.position = 'top',
            label.hide = FALSE,
            width = 10,
            color = 'gray',
            tile = 'OpenStreetMap.HOT',
            opacity = 0.4) |> 
  map.feature(languages = for_map$language,
            latitude = for_map$lat,
            longitude = for_map$lon,
            minichart.data = for_map |> select(-settlement, -lat, -lon, -language),
            minichart = 'pie', 
            width = 3,
            tile = 'OpenStreetMap.HOT',
            pipe.data = _)
}
```

### Data

```{r}
db |> 
  filter(feature_lexeme == 'Perfective stem') |> 
  select(settlement, value, stimuli, answer, collected) |> 
  DT::datatable(class = 'cell-border stripe', 
    rownames = FALSE, 
    filter = 'top', 
    extensions = 'Buttons',
    options = list(pageLength = 42, 
                   autoWidth = TRUE, 
                   info = FALSE,
                   dom = 'fBltp',
                   buttons = list(list(extend = 'collection',
                                       buttons = c('csv', 'excel', 'pdf'),
                                       text = '<i class="fas fa-download"></i>')),
                   paginate = TRUE))
```


