---
title: "Verb form used in present habitual contexts"
author: "Timur Maisak"
date: 2025-08-28
format: html
css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9.5)
```

The map shows the choice of verb forms in present habitual contexts ('He always cooks meat without salt', 'She milks the cow every evening'). Most varieties use the present habitual form, which is a periphrastic form consisting of the imperfective converb and the present-tense copula (IPFV.CVB + COP). In some varieties, the copula is omitted and the form is thus syncretic with the bare converb (IPFV.CVB). In some varieties, the converb marker -r / -rV is dropped and the lexical part is represented by the imperfective stem (IPFV + COP). Some varieties use the present tense in this context, which is a periphrastic form with another auxiliary verb (IPFV.CVB + 'be'). See also @FeatureX and @FeatureXX on the imperfective forms.


## Feature Map

::: {.panel-tabset}

### Map

```{r}
# Load data for this feature
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(lingtypology)

# Read the database and villages data
db <- read_csv("data/database.csv", show_col_types = FALSE)
villages <- read_csv("data/villages.csv", show_col_types = FALSE)

# Filter villages like in original script (exclude certain villages)
villages4map <- villages |> 
  filter(!(village %in% c("Kazankulak", "Novyy Borch", "Vrush", "Aran", "Khnyukh")))

# Create mapping data for this feature
feature_data <- db |> 
  filter(feature_id == 175) |>  # Filter by feature_id
  filter(!is.na(value), value != "NO DATA")

if (nrow(feature_data) > 0 && nrow(villages4map) > 0) {
  # Create mapping data for this feature
  mapping_data <- feature_data |>
    mutate(value = str_split(value, " ; ")) |> 
    unnest_longer(value) |> 
    distinct(settlement, value) |> 
    mutate(n = 1) |> 
    pivot_wider(names_from = value, values_from = n, values_fill = 0) |> 
    left_join(villages4map[,c("village","lat","lon")], by = c("settlement" = "village")) |> 
    mutate(language = "Rutul") |>
    filter(!is.na(lat) & !is.na(lon))
  
  if (nrow(mapping_data) > 0) {
    # Convert feature columns to numeric
    feature_cols <- setdiff(colnames(mapping_data), c("settlement","lat","lon","language"))
    mapping_data <- mapping_data |> mutate(across(all_of(feature_cols), as.numeric))
    
    # Create sophisticated mapping approach with ALWAYS visible village names
    if(length(feature_cols) == 1){
      # single feature column - show all villages in gray, feature data in color
      base_map <- map.feature(languages = "Rutul",
                             latitude = villages4map$lat,
                             longitude = villages4map$lon,
                             label = villages4map$village,
                             label.position = "top",
                             label.hide = FALSE,
                             width = 10,
                             color = "gray",
                             tile = "OpenStreetMap.HOT",
                             opacity = 0.4)
      
      feature_map <- map.feature(languages = "Rutul",
                                latitude = mapping_data$lat,
                                longitude = mapping_data$lon,
                                label = mapping_data$settlement,
                                label.position = "top",
                                label.hide = FALSE,
                                width = 10,
                                tile = "OpenStreetMap.HOT",
                                features = feature_cols)
      
      # Show both maps
      base_map
      feature_map
    } else {
      # multiple feature columns - use pie charts with ALWAYS visible village names
      base_map <- map.feature(languages = "Rutul",
                             latitude = villages4map$lat,
                             longitude = villages4map$lon,
                             label = villages4map$village,
                             label.position = "top",
                             label.hide = FALSE,
                             width = 10,
                             color = "gray",
                             tile = "OpenStreetMap.HOT",
                             opacity = 0.4)
      
      feature_map <- map.feature(languages = "Rutul",
                                latitude = mapping_data$lat,
                                longitude = mapping_data$lon,
                                minichart.data = mapping_data |> select(all_of(feature_cols)),
                                minichart = "pie",
                                width = 3,
                                tile = "OpenStreetMap.HOT",
                                label = mapping_data$settlement,
                                label.position = "top",
                                label.hide = FALSE)
      
      # Show both maps
      base_map
      feature_map
    }
  } else {
    cat("No geographical coordinates available for mapping.\n")
  }
} else {
  cat("No data available for this feature or no village coordinates found.\n")
}
```

### Data

```{r}
# Load required libraries
library(readr)
library(dplyr)
library(DT)

# Read the database data
db <- read_csv("data/database.csv", show_col_types = FALSE)

# Create interactive table for this feature
feature_data <- db |> 
  filter(feature_id == 175) |>  # Filter by feature_id
  filter(!is.na(value), value != "NO DATA")

if (nrow(feature_data) > 0) {
  # Prepare data for display
  display_data <- feature_data |>
    select(settlement, value, stimuli, answer, collected) |>
    arrange(settlement, value)
  
  # Use DT package for interactive table
  DT::datatable(display_data, 
                class = "cell-border stripe",
                rownames = FALSE,
                filter = "top",
                options = list(pageLength = 25, 
                             autoWidth = TRUE,
                             info = FALSE))
} else {
  cat("No data available for this feature.\n")
}
```

:::

