---
title: Choice of pronoun in reflexive contexts in 1st person
author: Maksim Melenchenko
date: 'Last update: `r lubridate::make_datetime(year = 2024, month = 3, day = 8)`'
output:
  html_document:
    number_sections: false
    anchor_sections: true
    pandoc_args: --shift-heading-level-by=-1
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE, fig.width = 9.5)
library(tidyverse)
library(lingtypology)
read_csv('data/database.csv', show_col_types = FALSE) |> 
  filter(feature_id == 191) ->
  db
read_csv('data/villages.csv') |> 
  filter(!(village %in% c('Kazankulak', 'Novyy Borch', 'Vrush', 'Aran', 'Khnyukh'))) ->
  villages
  
villages |> 
  filter(!(village %in% c('Tsudik', 'Borch'))) ->
  villages4map
  
```

This feature shows the choice of pronoun which is used in reflexive contexts in the 1st person: the 1st person pronoun or the reflexive pronoun. The speakers were asked to translate the following sentence: ‘**I**’ll go pour **myself** some tea’ (in Russian: ‘Пойду-ка **я** налью **себе** чаю’). For each speaker, first reaction was coded (although the elicitor tried to make sure that this reaction was not an error due to the misunderstanding of the context).
In all villages except Khnov, the 1st person pronoun (*zɨ[je]* / *zad* in absolutive, *zas* in dative) was used, while Khnov speakers used the reflexive pronoun.


##

::: {.panel-tabset}

### Map

```{r}
db |> 
  filter(!is.na(value),
         value != 'NO DATA') |>
  mutate(value = str_split(value, ' ; ')) |> 
  unnest_longer(value) |> 
  distinct(settlement, value) |> 
  mutate(n = 1) |> 
  pivot_wider(names_from = value, values_from = n, values_fill = 0) |> 
  left_join(villages[,c('village', 'lat', 'lon')], c('settlement' = 'village')) |> 
  mutate(language = 'Rutul') ->
  for_map
  
if(length(for_map) == 5){
map.feature(languages = 'Rutul',
            latitude = villages4map$lat,
            longitude = villages4map$lon,
            label = villages4map$village,
            label.position = 'top',
            label.hide = FALSE,
            width = 10,
            color = 'gray',
            tile = 'OpenStreetMap.HOT',
            opacity = 0.4) |> 
  map.feature(languages = for_map$language,
              latitude = for_map$lat,
              longitude = for_map$lon,
              label = for_map$settlement,
              label.position = 'top',
              label.hide = FALSE,
              width = 10,
              tile = 'OpenStreetMap.HOT',
              features = colnames(for_map)[2],
              pipe.data = _)  
} else {
map.feature(languages = 'Rutul',
            latitude = villages4map$lat,
            longitude = villages4map$lon,
            label = villages4map$village,
            label.position = 'top',
            label.hide = FALSE,
            width = 10,
            color = 'gray',
            tile = 'OpenStreetMap.HOT',
            opacity = 0.4) |> 
  map.feature(languages = for_map$language,
            latitude = for_map$lat,
            longitude = for_map$lon,
            minichart.data = for_map |> select(-settlement, -lat, -lon, -language),
            minichart = 'pie', 
            width = 3,
            tile = 'OpenStreetMap.HOT',
            pipe.data = _)
}
```

### Data

```{r}
db |> 
  select(settlement, value, stimuli, answer, collected) |> 
  DT::datatable(class = 'cell-border stripe', 
    rownames = FALSE, 
    filter = 'top', 
    extensions = 'Buttons',
    options = list(pageLength = 42, 
                   autoWidth = TRUE, 
                   info = FALSE,
                   dom = 'fBltp',
                   buttons = list(list(extend = 'collection',
                                       buttons = c('csv', 'excel', 'pdf'),
                                       text = '<i class="fas fa-download"></i>')),
                   paginate = TRUE))
```
:::

